---
title: "Global suicide Rates and its correlation to country income, region, age, and gender in 2019"
subtitle: "An analysis of global suicide rates in 2019 using multiple linear regression"
author: 
  - Hyuk Jang
thanks: "Code and data are available at: https://github.com/anggimude/Health."
date: today
date-format: long
abstract: "Suicide has become a main cause of deaths globally over the decades. This creates curiosity in what immutable factors may be affecting suicide rates. This paper looks into the correlation between suicide rates and gender, sex, age, country income, and region. Then further analysis is done by looking into human resources available in each country."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# work space setup
library(tidyverse)
library(dplyr)
library(arrow)
library(rstanarm)
library(ggplot2)
library(modelsummary)
library(tidybayes)
library(kableExtra)

# read data

sum_sta <- read_csv("~/Health/data/analysis_data/sum_sta.csv")
sum_sta_long <- read_csv("~/Health/data/analysis_data/sum_sta_long.csv")
both_table <- read_csv("~/Health/data/analysis_data/both_table.csv")

```

\newpage

# Introduction

Suicide is defined as the act of intentionally causing one's own death. The main risk factors of suicide are mental disorders, physical disorders, and substance abuse. In 2019, 1.3% of all deaths resulted from suicide and is the fourth leading cause of deaths in 15~29 years old. This paper will dive deeper into suicide rates and its correlation between country income, age, gender, and region in 2019. The most recent year with relevant data is for 2019, from the WHO[citation] and since we are only looking into one year, we look at the global trends and world bank income group trends to get a more thorough understanding of the trends of suicide rates of 2019. Because one of the goals of the paper is to observe the correlation between suicide rates and country income and region, the countries are classified based on the world bank income group; with four income groups - low, lower-middle, upper-middle, and high. Low-income economies is defined as a country with a gross national income less than \$1135, and in the manner lower-middle is between the range \$1136 to \$4465, upper-middle in the range of \$4455 to \$13845 and high is \$13846 or more. Region is defined based on the continents; Asia, Africa, North America, South America, and Europe. The total number of countries is 183 which are all mapped to different regions and income groups.   

This paper will produce a multiple linear regression to analyze the correlation coefficients of each factor. The results of the model will be checked by doing a 95% confident interval on the regression coefficients. In addition, we will be able to study the trends of the factors of suicide. We want to analyze the change in demographics for the different age groups, country income, region, and gender. The data was obtained from the WHO database[citaiton] under the mental health category. @citeR was used to clean the raw data to what we use for modelling in other to write an analysis of the intended study. The estimand of this paper are the intercepts and coefficients of each predictors which correspond to region/income/sex and the different age categories. 

This paper has 4 sections in total not including the introduction. In @sec-data we look at the data that used to carry out the reports including tables and graphs of cleaned data that will be used for the models and the summary statistics. In the next section, we discuss about the models that will be used to analyze our cleaned data, how it is set up and the justifications of it. Next we display and examine the results obtained from the models including tables of the model summaries which helps us make predictions. Lastly, we make final discussions of our results and research based on each cause and dive into some weaknesses that our paper has. In addition, we explore some next steps we or anyone else interested is willing to take after reading this paper. [make this paragraph more specific]


# Data {#sec-data}

## Raw Data

The data used in this paper is derived from WHO(@who2021global), and WHO data repository() and was downloaded from the data available in the links. WHO(@who2021global) provides data of age-standardized suicide rates per 100,000 population for 183 countries for the years 2000-2019. WHO data repository() provides data of crude suicide rates per 100,000 population for the same 183 countries for the year 2019. Since the only data avaiable data for crude suicide rates depending on age is 2019, we only look into the year 2019 for WHO() as well. The data the two tables include are the 183 countries, sex (both, male, female), age groups of 85+, 75-84, 65-74, 55-64, 45-54, 35-44, 25-34, 15-24, and age standardized suicide rates. The age-standardized suicide rate is used as the dependent variable for our model as we are interested in the correlation between the independent variables which are gender, age group, country income, and region.

The cleaning, testing, and modelling of the data for this paper was done through R [@citeR] with the aid of the following packages: tidyverse [@citetidyverse], dplyr [@citedplyr], rstanarm [@citerstanarm], ggplot2 [@citeggplot2], modelsummary [@citemodelsummary], kableExtra [@citekableExtra], arrow[@citearrow], tidybayes[@citetidybayes].

## Cleaned Data

The goal of the cleaning process of this paper is to create a table including income group(high, upper middle, lower middle, low), sex(both, male, female), region(Asia, Europe, North America, South America, Africa, Oceania) as the rows, and age-standardized suicide rates, and the 10 year age group suicide rates. This is so that we can create a model which will provide us the correlation of global suicide rates of 2019 against income group/sex/region and age groups. To do this we merge the two raw data tables we have downloaded. Before we merge the tables, WHO() includes the age-standardized suicide rates as a form of the actual value[lower limit, upper limit]. When we create our models or graphs, the interval is not necessary, so we first remove the interval so that each column only consists of the actual rate. After merging and doing some minor cleaning like removing repeated rows and changing names of columns and rows, the table will have country, sex, age-standardized suicide rates, and the 10 year age groups. We must also include the region and income group of each data so we create a mapping so that we country will have additional columns of continent and world bank income group. Since we are looking into the global suicide rates and how the independent variables affect it, we must calculate the averages of the age-standardized suicide rates and age groups depending on sex, income group, region. 



```{r}
#| label: tbl-sum_sta
#| tbl-cap: Global Suicide Rates by Age, Region, Sex, and Income Group
#| echo: false
#| warning: false

head(sum_sta, n=13) |>
  kable(
    col.names = c("Region/Income/Sex", "Age-standardized suicide rates (per 100 000 population)", "Age_85+", "Age_75-84", "Age_65-74", "Age_55-64", "Age_45-54", "Age_35-44", "Age_25-34", "Age_15-24"),
    align = c("l", "r", "r", "r", "r", "r", "r", "r", "r", "r"),
    digits = 2, 
    booktabs = TRUE, 
    linesep = ""
  ) 
```

```{r}
#| label: fig-sum_sta_long
#| fig-cap: lobal Suicide Rates by Age, Region, Sex, and Income Group
#| echo: false
#| warning: false
#| message: false

ggplot(sum_sta_long, aes(x = `Region/Income/Sex`, y = Suicide_Rate, color = Age_Group)) +
  geom_point() +
  ggtitle("Global Suicide Rates by Age, Region, Sex, and Income Group") +
  xlab("Region/Income Group/Sex and Age Group") +
  ylab("Suicide Rate") +
  theme_minimal()
```


## Basic Summary Statistics

@tbl-sumsta_korea is a representation of the @tbl-intersect_data_one showing its minimum, mean, maximum, standard deviation, variance, and sample size. The summary statistics shows that the mean of the number of deaths are 18320, with a minimum of 6860, and a maximum of 44109 for a sample size of 24 as the data selected is from @tbl-intersect_data_one. The standard deviation and variance may be abnormally high because the range of the data is large, in other words, because the mean is 18320, and the maximum is 44109, there is likely a few outliers in the data creating such a high standard deviation and variance.

\FloatBarrier
```{r}
#| echo: false
#| eval: true
#| label: tbl-sumsta_suicide
#| tbl-cap: "Summary statistics of the number of yearly deaths, by cause, in South Korea"


datasummary(
  (`Age-standardized suicide rates (per 100 000 population)` + `Age_85+` + `Age_75-84` + `Age_65-74` + `Age_55-64` + `Age_45-54` + `Age_35-44` + `Age_25-34` + `Age_15-24`) ~
    Min + Mean + Max + SD + Var + N,
  fmt = 0,
  data = sum_sta
)

```

```{r}
#| label: tbl-avg_data
#| tbl-cap: Average global data
#| echo: false
#| warning: false

head(average_data) |>
  kable(
    col.names = c("Region/Income/Sex", "Age-standardized suicide rates (per 100 000 population)", "Age_85+", "Age_75-84", "Age_65-74", "Age_55-64", "Age_45-54", "Age_35-44", "Age_25-34", "Age_15-24"),
    align = c("l", "r", "r", "r", "r", "r", "r", "r", "r", "r"),
    digits = 2, 
    booktabs = TRUE, 
    linesep = ""
  ) 
```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


